---
title: "Data Import"
output: github_document
---

```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)
library(readxl)
library(haven)
```

This document will show how to import data.


## Import the FAS Litters CSV
```{r}
litters_df = read_csv('data/FAS_litters.csv')

litters_df = janitor::clean_names(litters_df) 
# clean_names(): space -> underscore, uppercase -> lowercase
# `::`: specifies which package the function is from; use if only one function is needed from the package
```


## Look at the dataset
```{r}
litters_df

head(litters_df)
tail(litters_df, 10)
```

```{r eval = FALSE}
# eval = FALSE helps with view() when knitting the document
view(litters_df)
```

## Learning assessment
Importing FAS_pups.csv using relative path

```{r}
pups_df_rel = read_csv('data/FAS_pups.csv')
pups_df_rel = janitor::clean_names(pups_df_rel)
pups_df_rel
```
Importing FAS_pups.csv using absolute path. Absolute paths are fragile and not sustainable between collaborators.
```{r}
pups_df_abs = read_csv('~/columbia/bist p8150/data_wrangling_i/data/FAS_pups.csv')
```

## Look at read_csv options
Importing data.

If data contains irrelevant info in the first few rows, use the `skip` argument.

```{r}
litters_df = read_csv(file = 'data/FAS_litters.csv',
                      col_names = FALSE)
# 
```

Working with missing data.

Use the `na` argument to replace all missing data and make sure data are numerical
```{r}
litters_df = read_csv(file = 'data/FAS_litters.csv',
                      na = c('NA', '', '.'))
```
If we want to change the data type of a column, use `cols()`
```{r}
litters_df = read_csv(file = 'data/FAS_litters.csv',
                      na = c('NA', '', '.'),
                      col_types = cols(Group = col_factor()))
litters_df
```

## Import an excel file

Importing the mlb11 excel file.

`read_excel()` is under the `readxl` package. 
Use the `sheet` argument to load in specific sheet in an excel.
```{r}
mlb_df = read_excel('data/mlb11.xlsx')
```

## Import SAS data
```{r}
pulse_df = read_sas('data/public_pulse_data.sas7bdat')
```

