---
title: "Data Manipulation"
output: github_document
---

```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)
```

This document will show how to *manipulate* data.

Import the 2 datasets that we're going to manipulate.
```{r message = FALSE}
litters_df = read_csv("data/FAS_litters.csv",
                      na = c("NA", "", "."))
litters_df = janitor::clean_names(litters_df)

pups_df = read_csv("data/FAS_pups.csv",
                   na = c("NA", "", "."))
pups_df = janitor::clean_names(pups_df)
```
## `select`

Use `select()` to select variables. First argument is always the df. 
Then inpt columns names 
```{r}
select(litters_df,
       group,
       litter_number,
       gd0_weight)
```

```{r}
select(litters_df, 
       group:gd18_weight)
```

```{r}
select(litters_df,
       -pups_survive)
```

```{r}
select(litters_df, 
       -(group:gd18_weight))
```

```{r}
select(litters_df,
       starts_with("gd"))

select(litters_df,
       contains("pups"))
```
Renaming column

```{r}
select(litters_df,
       GROUP = group)

rename(litters_df, 
       GROUP = group)
```

Relocating columns

```{r}
select(litters_df,
       litter_number,
       gd0_weight,
       everything())

relocate(litters_df, 
         litter_number, 
         gd0_weight)
```

## Learning assessment
```{r}
select(pups_df, litter_number, sex, pd_ears)
```

## `filter`

Filtering out rows.

```{r}
filter(litters_df,
       gd_of_birth == 20)
```

```{r}
filter(litters_df,
       pups_born_alive > 8)
```

```{r}
filter(litters_df,
       pups_born_alive != 9)
```

Use `%in%` if more than one rows (OR gate)

```{r}
filter(litters_df,
       group %in% c("Low8", "Low7"))
```

```{r}
# AND gate
filter(litters_df,
       group %in% c("Low8", "Low7"),
       pups_born_alive == 8)

# OR gate
filter(litters_df,
       group %in% c("Low8", "Low7") | pups_born_alive == 8)
```
```{r}
# drop all NA
drop_na(litters_df)

# drop NA in a specific column
drop_na(litters_df, gd0_weight)
```

## Learning assessment
```{r}
filter(pups_df,
       sex == 1)

filter(pups_df,
       pd_walk < 11,
       sex == 2)
```

## `mutate`

Can mutate values and strings in column. New columns append on the right.

```{r}
mutate(litters_df,
       wt_gain = gd18_weight - gd0_weight)

mutate(litters_df,
       group = str_to_lower(group))
```

Mutate multiple columns

```{r}
mutate(litters_df,
       wt_gain = gd18_weight - gd0_weight,
       group = str_to_lower(group))
```

## `arrange`

Sort values in a column (default = low to high). Can also sort based on multiple columns

```{r}
arrange(litters_df,
        gd0_weight)

arrange(litters_df,
        desc(gd0_weight))

arrange(litters_df,
        pups_born_alive,
        gd0_weight)
```

